### 2017-03-31 ThesisText/experiments.tex

\section{Students, Courses, Logs, and Results}
\label{sec:exp_001_002}

@@

\[
    [0, 10[\;,\, [10, 12[\;,\, [12, 14[\;,\, [14, 16[\;,\, [16, 18[\;,\, [18, 20]
\]

@@

\[
    [0, 10[\;,\, [10, 14[\;,\, [14, 18[\;,\, [18, 20]
\]

@@

\subsection{Predicting Students Grades Based on Moodle Usage}
% EXP_001

A first experiment on this dataset aims at predicting the students grades based
on their Moodle usage data and in their other characteristics present on this
dataset. In this experiment six tasks are made. Both Decision Trees and Bayesian
Networks are trained using the dataset with binary classes, classes with
intervals of two, and classes with intervals of four.

Some fields on dataset~\ref{tab:dat_001} are strings and enumerated types,
therefore, these are nominal fields. Because of this characteristic, Decision
Trees and Bayesian Networks are used.

@@

    \begin{tabular}{| l | l | r | r | r |}
        \hline
        \textbf{Model} & \textbf{Classes} & \textbf{Maximum} & \textbf{Minimum} & \textbf{Average} \\ \hline
        Decision Tree    & Binary Class          & 0.92 & 0.34 & 0.70 \\
                         & Grades Classes 4 by 4 & 0.84 & 0.22 & 0.59 \\
                         & Grades Classes 2 by 2 & 0.74 & 0.34 & 0.55 \\ \hline
        Bayesian Network & Binary Class          & 0.95 & 0.53 & 0.74 \\
                         & Grades Classes 4 by 4 & 0.73 & 0.36 & 0.57 \\
                         & Grades Classes 2 by 2 & 0.67 & 0.28 & 0.52 \\ \hline
    \end{tabular}

@@

It is observable that for binary classes, Bayesian Networks performed slightly
better then decision trees. But for grade classes two by two and four by four,
decision trees were better. In particular, training models for binary classes,
understandably, yields better results on average.

@@

\subsection{Clustering Students Grades Based on Moodle Usage}
% EXP_002

@@

    \caption
        []
        {}

@@

\section{Number of Courses and Logs}

@@

Some classification tasks use the \texttt{ApprovedCourses} feature as a class. In this case, feature may be left in it's original form or it may be replaced by classes. There are three ways to do this replacement.~\TODO{Explain this and mentioned statistics}

\subsection{Predicting Number of Approved Courses Based on Moodle Usage}
% EXP_003

Like before both Decision Trees and Bayesian Networks are used.
Table~\ref{tab:exp_003_res} shows the maximum, minimum, and average accuracy
for each pair of model and classes.

@@

    \begin{tabular}{| l | l | r | r | r |}
        \hline
        \textbf{Model} & \textbf{Classes} & \textbf{Maximum} & \textbf{Minimum} & \textbf{Average} \\ \hline
        Decision Tree    & 2 Classes        & 1.00 & 0.64 & 0.83 \\
                         & 4 Classes        & 1.00 & 0.29 & 0.68 \\
                         & 6 Classes        & 1.00 & 0.21 & 0.68 \\
                         & Original Classes & 0.86 & 0.14 & 0.52 \\ \hline
        Bayesian Network & 2 Classes        & 1.00 & 0.47 & 0.70 \\
                         & 4 Classes        & 0.87 & 0.29 & 0.56 \\
                         & 6 Classes        & 0.86 & 0.07 & 0.43 \\
                         & Original Classes & 0.86 & 0.00 & 0.39 \\ \hline
    \end{tabular}

@@

Bayesian Networks are achieve hight level of accuracy, but not as high as
Decision Trees. For the dataset with the original classes, Bayesian Networks
perform similarly in the best case scenario, given that the values for maximum
accuracy are similar. But the values for minimum and average are noticeably
lower.

@@

\subsection{Clustering Number of Approved Courses Based on Moodle Usage}
% EXP_004

@@

\section{Student/Course Profile and Results}

@@

\subsection{Predicting Students Grades From Student/Course Profiling}
% EXP_005

@@

    \begin{tabular}{| l | l | r | r | r |}
        \hline
        \textbf{Model} & \textbf{Classes} & \textbf{Maximum} & \textbf{Minimum} & \textbf{Average} \\ \hline
        Decision Tree    & Binary Class          & 0.91 & 0.43 & 0.73 \\
                         & Grades Classes 4 by 4 & 0.65 & 0.31 & 0.44 \\
                         & Grades Classes 2 by 2 & 0.56 & 0.18 & 0.34 \\ \hline
        Bayesian Network & Binary Class          & 0.97 & 0.49 & 0.77 \\
                         & Grades Classes 4 by 4 & 0.91 & 0.31 & 0.55 \\
                         & Grades Classes 2 by 2 & 0.71 & 0.26 & 0.46 \\ \hline
    \end{tabular}

@@

We observe from the results that Bayesian Networks perform better then Decision
Trees achieving greater levels of accuracy for each dataset variant.

\subsection{Clustering Students Grades From Student/Course Profiling}
% EXP_006

@@

    \caption
        []
        {}

>>>>

% =============================================================================
\section{Students, Courses, Logs, and Results}
\label{sec:exp_001_002}

@@

\[
    [0, 10[\;,\, [10, 12[\;,\, [12, 14[\;,\, [14, 16[\;,\, [16, 18[\;,\, [18, 20].
\]

@@

\[
    [0, 10[\;,\, [10, 14[\;,\, [14, 18[\;,\, [18, 20].
\]

@@

\subsection{Predicting Students Grades Based on Moodle Usage}
\label{sec:exp_001}
% EXP_001

A first experiment on this dataset aims at predicting the students grades based
on their Moodle usage data and in their other characteristics present on this
dataset. In this experiment six tasks are made. Both Decision Trees and Naive
Bayes Classifiers are trained using the dataset with binary classes, classes
with intervals of two, and classes with intervals of four.

Some fields on dataset~\ref{tab:dat_001} are strings and enumerated types,
therefore, these are nominal fields. Because of this characteristic, Decision
Trees and Naive Bayes Classifiers are used.

@@

    \begin{tabular}{| l | l | r | r | r |}
        \hline
        \textbf{Model} & \textbf{Classes} & \textbf{Maximum} & \textbf{Minimum} & \textbf{Average} \\ \hline
        Decision Tree & Binary Class          & 0.92 & 0.34 & 0.70 \\
                      & Grades Classes 4 by 4 & 0.84 & 0.22 & 0.59 \\
                      & Grades Classes 2 by 2 & 0.74 & 0.34 & 0.55 \\ \hline
        Naive Bayes   & Binary Class          & 0.95 & 0.53 & 0.74 \\
                      & Grades Classes 4 by 4 & 0.73 & 0.36 & 0.57 \\
                      & Grades Classes 2 by 2 & 0.67 & 0.28 & 0.52 \\ \hline
    \end{tabular}

@@

It is observable that for binary classes, Naive Bayes Classifiers performed
slightly better then decision trees. But for grade classes two by two and four
by four, decision trees were better. In particular, training models for binary
classes, understandably, yields better results on average.

@@

\subsection{Clustering Students Grades Based on Moodle Usage}
\label{sec:exp_002}
% EXP_002

@@

    \caption
        [Clustering Students Grades Based on Moodle Usage]
        {Clustering Students Grades Based on Moodle Usage. The table displays
        the results for the silhouette score for clustering. The values for
        every variant of the dataset are approximately the same, so only one
        table is shown.}

@@

% =============================================================================
\section{Number of Courses and Logs}
\label{sec:exp_003_004}

@@

Some classification tasks use the \texttt{ApprovedCourses} feature as a class.
In this case, feature may be left in it's original form or it may be replaced
by classes. There are three ways to do this replacement.

The maximum number of approved courses is 12 and minimum is 0, so the possible
divisions are with two classes:

\[
    [0, 6[\;,\, [6, 12],
\]

Three classes:

\[
    [0, 4[\;,\, [4, 8[\;,\, [8, 12],
\]

And four classes:

\[
    [0, 3[\;,\, [3, 6[\;,\, [6, 9[\;,\, [9, 12].
\]

\subsection{Predicting Number of Approved Courses Based on Moodle Usage}
\label{sec:exp_003}
% EXP_003

Like before both Decision Trees and Naive Bayes Classifiers are used.
Table~\ref{tab:exp_003_res} shows the maximum, minimum, and average accuracy
for each pair of model and classes.

@@

    \begin{tabular}{| l | l | r | r | r |}
        \hline
        \textbf{Model} & \textbf{Classes} & \textbf{Maximum} & \textbf{Minimum} & \textbf{Average} \\ \hline
        Decision Tree & 2 Classes        & 1.00 & 0.64 & 0.83 \\
                      & 4 Classes        & 1.00 & 0.29 & 0.68 \\
                      & 6 Classes        & 1.00 & 0.21 & 0.68 \\
                      & Original Classes & 0.86 & 0.14 & 0.52 \\ \hline
        Naive Bayes   & 2 Classes        & 1.00 & 0.47 & 0.70 \\
                      & 4 Classes        & 0.87 & 0.29 & 0.56 \\
                      & 6 Classes        & 0.86 & 0.07 & 0.43 \\
                      & Original Classes & 0.86 & 0.00 & 0.39 \\ \hline
    \end{tabular}

@@

Naive Bayes Classifiers are achieve hight level of accuracy, but not as high as
Decision Trees. For the dataset with the original classes, Naive Bayes
Classifiers perform similarly in the best case scenario, given that the values
for maximum accuracy are similar. But the values for minimum and average are
noticeably lower.

@@

\subsection{Clustering Number of Approved Courses Based on Moodle Usage}
\label{sec:exp_004}
% EXP_004

@@

% =============================================================================
\section{Student/Course Profile and Results}
\label{sec:exp_005_006}

@@

\subsection{Predicting Students Grades From Student/Course Profiling}
\label{sec:exp_005}
% EXP_005

@@

    \begin{tabular}{| l | l | r | r | r |}
        \hline
        \textbf{Model} & \textbf{Classes} & \textbf{Maximum} & \textbf{Minimum} & \textbf{Average} \\ \hline
        Decision Tree & Binary Class          & 0.91 & 0.43 & 0.73 \\
                      & Grades Classes 4 by 4 & 0.65 & 0.31 & 0.44 \\
                      & Grades Classes 2 by 2 & 0.56 & 0.18 & 0.34 \\ \hline
        Naive Bayes   & Binary Class          & 0.97 & 0.49 & 0.77 \\
                      & Grades Classes 4 by 4 & 0.91 & 0.31 & 0.55 \\
                      & Grades Classes 2 by 2 & 0.71 & 0.26 & 0.46 \\ \hline
    \end{tabular}

@@

We observe from the results that Naive Bayes Classifiers perform better then
Decision Trees achieving greater levels of accuracy for each dataset variant.

\subsection{Clustering Students Grades From Student/Course Profiling}
\label{sec:exp_006}
% EXP_006

@@

    \caption
        [Clustering Students Grades From Student/Course Profiling]
        {Clustering Students Grades From Student/Course Profiling. The values
        for every variant of the dataset are approximately the same, so only
        one table is shown.}

### 2017-03-31 ThesisText/final.tex

\begin{enumerate}
    \item Predict student grades from their usage of Moodle;
        (Section~\ref{sec:exp_001})
    \item Predict how many courses is a student going to complete given he's
        usage of Moodle and initial number of enrolled courses;
        (Section~\ref{sec:exp_003})
    \item Predict student grades from his usage of features in Moodle.
        (Section~\ref{sec:exp_005})
\end{enumerate}

>>>>

\begin{enumerate}
    \item Predict student grades from their usage of Moodle;
        (Section~\ref{sec:exp_001})
    \item Predict how many courses a student is going to complete given his
        usage of Moodle and initial number of enrolled courses;
        (Section~\ref{sec:exp_003})
    \item Predict student grades from his usage of features in Moodle.
        (Section~\ref{sec:exp_005})
\end{enumerate}

### 2017-03-31 ThesisText/introduction.tex

These E-Leaning platforms are called Learning Management Systems. The platforms
allow institutions not only to deliver training but also keep track of users,
courses, content, etc. In their normal operations, theses platforms generate
and store a lot of data. This data concerns the courses, the content in those
courses, activities from users, and so on. In some specific cases, like Moodle
in universities, the system may even keep track of the students grades.

@@

The field of study which deals with applying Data Mining techniques to
educational data is called Educational Data Mining. This projects applies EDM
to data from the Moodle of the University of Évora.~\cite{ind_010, ind_011,
ind_013, ind_014, ind_015}

@@

Some statistical analysis are made to the pre processed data repository before
any Data Mining techniques are applied. This is done in order to to understand
the contents of the data and know which~\gls{dm} techniques should be applied
and to which subset of the data repository those techniques should be applied
to. Section~\ref{sec:stats} talks about these statistics and draws some
conclusions from them.

>>>>

These E-Leaning platforms are called~\gls{lms}. The platforms
allow institutions not only to deliver training but also keep track of users,
courses, content, etc. In their normal operations, theses platforms generate
and store a lot of data. This data concerns the courses, the content in those
courses, activities from users, and so on. In some specific cases, like Moodle
in universities, the system may even keep track of the students grades.

@@

The field of study which deals with applying~\gls{dm} techniques to educational
data is called~\gls{edm}. This projects applies~\gls{edm} to data from the
Moodle of the University of Évora.~\cite{ind_010, ind_011, ind_013, ind_014,
ind_015}

@@

Some statistical analysis are made to the pre processed data repository before
any~\gls{dm} techniques are applied. This is done in order to to understand the
contents of the data and know which~\gls{dm} techniques should be applied and
to which subset of the data repository those techniques should be applied to.
Section~\ref{sec:stats} talks about these statistics and draws some conclusions
from them.
